---
title: "Sales"
author: "Ethan Chan, Alex Back, Darren Kuboyama"
date: "12/11/2024"
format:
  html:
    toc: true
    toc-location: left
    embed-resources: true
    code-line-numbers: true
---

**Scientific Question: What differentiates top-rated products from bottom-rated products across various categories, and what keywords in customer reviews signal positive sentiments?**

We begin by adding the tidyverse library for later uses.

#Packages

```{r}
library(tidyverse)
library(ggplot2)
```

Next we read in the product information into a dataframe and then combine all the review csvs into one single dataframe to later data clean.

```{r}
product_info <- read.csv('product_info.csv')
review_files <- list.files(pattern = "reviews_.*\\.csv") #merge csvs
reviews <- do.call(rbind, lapply(review_files, read.csv)) #make it into a dataframe
```

Now we narrow it down to the metrics we want to look at, specifically the categories, reviews, ratings, and the love counts.

```{r}
product_data <- product_info %>%
  select(product_id, product_name, brand_id, secondary_category, tertiary_category, rating, reviews, loves_count) %>%
  filter(!(secondary_category == "" & tertiary_category == ""))
```

Now we sort them into the top 3 and worst 3 based on the secondary category.

```{r}
top_3 <- product_data %>%
  group_by(secondary_category) %>%
  arrange(desc(rating), desc(reviews), desc(loves_count)) %>% 
  slice_head(n = 3) %>%  
  ungroup()

worst_3 <- product_data %>%
  group_by(secondary_category) %>%
  arrange(rating, reviews, loves_count) %>%  
  slice_head(n = 3) %>%  
  ungroup()

combined <- bind_rows(
  top_3 %>% mutate(category = "Top 3"),
  worst_3 %>% mutate(category = "Worst 3")
)

```

Now we might look at some trends, lets look at the average rating, reviews, and loves of the top 3 and worst 3.

```{r}
# Combined summary for Top 3 and Worst 3
summary_overall <- combined %>%
  group_by(category) %>%
  summarize(
    avg_rating = mean(rating, na.rm = TRUE),
    avg_reviews = mean(reviews, na.rm = TRUE),
    avg_loves = mean(loves_count, na.rm = TRUE)
  )
summary_category <- combined %>%
  group_by(secondary_category, category) %>%
  summarize(
    avg_rating = mean(rating, na.rm = TRUE),
    avg_reviews = mean(reviews, na.rm = TRUE),
    avg_loves = mean(loves_count, na.rm = TRUE),
    .groups = "drop"
  )
```

We can now start comparing the rating, reviews, and love counts for the top 3 and bottom 3 products of each category by looking at some graphs.

#Graphs

```{r}
top_bottom_moisturizers <- combined %>%
  filter(secondary_category == "Moisturizers") %>%
  filter(category %in% c("Top 3", "Worst 3"))  

print(top_bottom_moisturizers)

ggplot(top_bottom_moisturizers, aes(x = reorder(product_name, rating), y = rating, fill = category)) +
  geom_bar(stat = "identity", width = 0.5) +
  labs(
    title = "Top 3 and Bottom 3 Ratings for Moisturizers",
    x = "Product Name",
    y = "Rating"
  ) +
  coord_flip() +  
  theme_minimal() +
  theme(
    axis.text.y = element_text(size = 6),  
    axis.text.x = element_text(size = 6)
  )
```
An important detail we can see just from visualizing the product names in this graph is that popular ingredients like Hyaluronic Acid does not guarantee a high rating. In fact, it is in the Worst 3 category. In contrast ingredients like Niacinamide or products that target dark spots and the skin's texture seem to be dominating the Top 3.

Now lets narrow this down to gain a deeper understanding of the metric's correlations. This will give us a deeper understanding of whether or not the high ratings are linked to customer engagement and the relationship of these metrics could be a strength or weakness in a company's product. 

```{r}
# Scatterplot: Rating vs Reviews
ggplot(combined, aes(x = reviews, y = rating)) +
  geom_point(alpha = 0.6, color = "blue") +
  geom_smooth(method = "lm", se = FALSE, color = "red") +
  labs(
    title = "Relationship Between Reviews and Rating",
    x = "Number of Reviews",
    y = "Rating"
  ) +
  theme_minimal()

# Scatterplot: Rating vs Loves Count
ggplot(combined, aes(x = loves_count, y = rating)) +
  geom_point(alpha = 0.6, color = "green") +
  geom_smooth(method = "lm", se = FALSE, color = "red") +
  labs(
    title = "Relationship Between Loves Count and Rating",
    x = "Loves Count",
    y = "Rating"
  ) +
  theme_minimal()

# Scatterplot: Loves Count vs Reviews
ggplot(combined, aes(x = reviews, y = loves_count)) +
  geom_point(alpha = 0.6, color = "purple") +
  geom_smooth(method = "lm", se = FALSE, color = "red") +
  labs(
    title = "Relationship Between Reviews and Loves Count",
    x = "Number of Reviews",
    y = "Loves Count"
  ) +
  theme_minimal()


```
The first graph, showing the relationship between reviews and ratings, also 
exhibits a weak positive correlation. Products with more reviews tend to have 
slightly higher ratings, but most ratings remain in the 4 to 5 range, regardless
of review count. Reviews highlight engagement and visibility, while ratings 
reflect customer satisfaction. The clustering of ratings indicates that while 
customer satisfaction is generally high, reviews are a more dynamic metric for 
measuring product visibility and interaction.

The second graph, focusing on the relationship between loves count and ratings, 
reveals a weak positive correlation. While products with higher loves tend to 
have slightly higher ratings, the relationship is not strong, and ratings 
cluster around 4 to 5 regardless of popularity. Loves count measures customer 
loyalty and emotional appeal, while ratings assess satisfaction. The lack of a 
strong correlation suggests that a product's popularity does not always 
translate directly into higher satisfaction. This emphasizes the importance of 
analyzing these metrics together to capture a product’s full performance.

The third graph, examining the relationship between reviews and loves count,
shows a strong positive correlation, indicating that products with higher 
reviews also tend to have higher loves. This suggests that customer interaction 
through reviews aligns closely with emotional resonance and loyalty, as measured
by loves count. Loves count reflects a product's desirability and brand appeal, 
while reviews showcase customer engagement and feedback. Products that perform 
well in both metrics demonstrate strong marketing and customer alignment.

Quantitatively, the correlation between reviews and loves is significantly 
stronger compared to the relationships between reviews and ratings or ratings 
and loves. This suggests that products with higher reviews are not only engaging
customers but are also more likely to be loved, making the combination of these 
metrics a holistic indicator of product performance.

```{r}
correlations <- combined %>%
  group_by(secondary_category) %>%
  summarize(
    corr_reviews_rating = cor(reviews, rating, use = "complete.obs"),  
    corr_rating_loves = cor(rating, loves_count, use = "complete.obs"),
    corr_reviews_loves = cor(reviews, loves_count, use = "complete.obs"), 
    .groups = "drop"
  )

correlations


```

This table reveals something interesting. The worst 3 products tend to have a higher average review and love count compared to the top 3 products which have fewer reviews and love counts. This suggests that high visibility and customer engagement (throguh reviews) dont always translate to positive sentiments or high ratings. instead these metrics could indicate polarizing products that attract significant attention but also receive miexed or negative feedback. However, the top 3 products,despite having fewer reviews and loves maintain a higher average rating which could mean they deliver consistent qualtiy and satisifcation to a samller but liekly more loyal audience. This observation could enforce the idea taht quantity of engagement throguh reviews or loves does not necessairly equate to a quality of sentiment through ratings. Understanding this dynamic could help a brand focus on improving their underperformign product with high visibility to turn into loyalty and satsifcation. But why might a low rated product have such low love counts you might ask? This could be due to either strong marketing, or a brand's reputation, a subset of users who strongly dislike and those who like, historical popularity that has declined, brand loyalty, or just simply an intent to purchase but never did.

```{r}
summary_metrics <- combined %>%
  group_by(category) %>%
  summarize(
    avg_rating = mean(rating, na.rm = TRUE),
    avg_reviews = mean(reviews, na.rm = TRUE),
    avg_loves = mean(loves_count, na.rm = TRUE),
    .groups = "drop"
  )

print(summary_metrics)

ggplot(summary_metrics, aes(x = category, y = avg_rating, fill = category)) +
  geom_bar(stat = "identity") +
  labs(
    title = "Average Rating Across Categories (Top 3 vs Worst 3)",
    x = "Category",
    y = "Average Rating"
  ) +
  theme_minimal() 

ggplot(summary_metrics, aes(x = category, y = avg_reviews, fill = category)) +
  geom_bar(stat = "identity") +
  labs(
    title = "Average Reviews Across Categories (Top 3 vs Worst 3)",
    x = "Category",
    y = "Average Reviews"
  ) +
  theme_minimal() 

ggplot(summary_metrics, aes(x = category, y = avg_loves, fill = category)) +
  geom_bar(stat = "identity") +
  labs(
    title = "Average Loves Across Categories (Top 3 vs Worst 3)",
    x = "Category",
    y = "Average Loves"
  ) +
  theme_minimal() 

```

Now to better understand our products by analyzing their performance through primary categories which provides a simplified and clearer view of data. However, to achieve this, we exclude the mini sizes, as they represent smaller versions of existing products and may skew the results without offering anything unique. We group these primary categories again into their ratings, reviews, and love counts, for the top 3 and bottom 3 products in each category. Skincare has higher reviews and love counts for the worst 3 products compared to the top 3 which could suggest high customer interest but mixed satisfaction or the fact that there were more people who bought the product but just didn't like it compared to the top 3. While tools and brushes excel in all the categories for the top 3 which might highlight a category where quality plays a critical role in customer rating. While the rest might just suggest niche appeals or polarizing customer preferences, these patterns emphasize that high engagement does not always translate to high satisfaction, as we've seen before. This underscores the importance that we must dive deeper into brand performance review sentiments to truly get the crux of it and uncover reasons behind these trends and understand strategies for product improvement.

```{r}
#use primary category now for graph simplification
combined_with_primary <- combined %>%
  left_join(product_info %>% select(product_id, primary_category), by = "product_id")

#remove "mini size" irrelevant and skews info
broad_category_summary <- combined_with_primary %>%
  filter(primary_category != "Mini Size") %>% 
  mutate(
    broad_category = ifelse(is.na(primary_category) | primary_category == "", "Other", primary_category)  
  ) %>%
  group_by(broad_category, category) %>%  
  summarize(
    avg_rating = mean(rating, na.rm = TRUE),   
    avg_reviews = mean(reviews, na.rm = TRUE), 
    avg_loves = mean(loves_count, na.rm = TRUE), 
    .groups = "drop"
  )

#avg rating
ggplot(broad_category_summary, aes(x = reorder(broad_category, avg_rating), y = avg_rating, fill = category)) +
  geom_bar(stat = "identity", position = "dodge") +
  coord_flip() +
  labs(
    title = "Average Ratings by Primary Categories (Top 3 vs Worst 3)",
    x = "Primary Category",
    y = "Average Rating",
    fill = "Category"
  ) +
  theme_minimal()

#avg review
ggplot(broad_category_summary, aes(x = reorder(broad_category, avg_reviews), y = avg_reviews, fill = category)) +
  geom_bar(stat = "identity", position = "dodge") +
  coord_flip() +
  labs(
    title = "Average Reviews by Primary Categories (Top 3 vs Worst 3)",
    x = "Primary Category",
    y = "Average Reviews",
    fill = "Category"
  ) +
  theme_minimal()

#avg love count
ggplot(broad_category_summary, aes(x = reorder(broad_category, avg_loves), y = avg_loves, fill = category)) +
  geom_bar(stat = "identity", position = "dodge") +
  coord_flip() +
  labs(
    title = "Average Loves Count by Primary Categories (Top 3 vs Worst 3)",
    x = "Primary Category",
    y = "Average Loves Count",
    fill = "Category"
  ) +
  theme_minimal()

```








