---
title: "Sales"
author: "Ethan Chan, Alex Back, Darren Kuboyama"
date: "12/11/2024"
format:
  html:
    toc: true
    toc-location: left
    embed-resources: true
    code-line-numbers: true
---

**Scientific Question: What differentiates top-rated products from bottom-rated products across various categories, and what keywords in customer reviews signal positive sentiments?**

We begin by adding the tidyverse library for later uses.

#Packages

```{r}
library(tidyverse)
library(ggplot2)
```

Next we read in the product information into a dataframe and then combine all the review csvs into one single dataframe to later data clean.

```{r}
product_info <- read.csv('product_info.csv')
review_files <- list.files(pattern = "reviews_.*\\.csv") #merge csvs
reviews <- do.call(rbind, lapply(review_files, read.csv)) #make it into a dataframe
```

Now we narrow it down to the metrics we want to look at, specifically the categories, reviews, ratings, and the love counts.

```{r}
product_data <- product_info %>%
  select(product_id, product_name, brand_id, secondary_category, tertiary_category, rating, reviews, loves_count) %>%
  filter(!(secondary_category == "" & tertiary_category == ""))
```

Now we sort them into the top 3 and worst 3 based on the secondary category.

```{r}
top_3 <- product_data %>%
  group_by(secondary_category) %>%
  arrange(desc(rating), desc(reviews), desc(loves_count)) %>% 
  slice_head(n = 3) %>%  
  ungroup()

worst_3 <- product_data %>%
  group_by(secondary_category) %>%
  arrange(rating, reviews, loves_count) %>%  
  slice_head(n = 3) %>%  
  ungroup()

combined <- bind_rows(
  top_3 %>% mutate(category = "Top 3"),
  worst_3 %>% mutate(category = "Worst 3")
)

```

Now we might look at some trends, lets look at the average rating, reviews, and loves of the top 3 and worst 3.

```{r}
# Combined summary for Top 3 and Worst 3
summary_overall <- combined %>%
  group_by(category) %>%
  summarize(
    avg_rating = mean(rating, na.rm = TRUE),
    avg_reviews = mean(reviews, na.rm = TRUE),
    avg_loves = mean(loves_count, na.rm = TRUE)
  )
summary_category <- combined %>%
  group_by(secondary_category, category) %>%
  summarize(
    avg_rating = mean(rating, na.rm = TRUE),
    avg_reviews = mean(reviews, na.rm = TRUE),
    avg_loves = mean(loves_count, na.rm = TRUE),
    .groups = "drop"
  )
```

We can now start comparing the rating, reviews, and love counts for the top 3 and bottom 3 products of each category by looking at some graphs.

#Graphs

```{r}
top_bottom_moisturizers <- combined %>%
  filter(secondary_category == "Moisturizers") %>%
  filter(category %in% c("Top 3", "Worst 3"))  # Grab Top 3 and Worst 3

#print(top_bottom_moisturizers) Testing that the correct products are printed

ggplot(top_bottom_moisturizers, aes(x = reorder(product_name, rating), y = rating, fill = category)) +
  geom_bar(stat = "identity", width = 0.5) +
  labs(
    title = "Top 3 and Bottom 3 Ratings for Moisturizers",
    x = "Product Name",
    y = "Rating"
  ) +
  coord_flip() +  # flip the axis for better visibility
  theme_minimal() +
  theme(
    axis.text.y = element_text(size = 6),  
    axis.text.x = element_text(size = 6)
  )
```
An important detail we can see just from visualizing the product names in this graph is that popular ingredients like Hyaluronic Acid does not guarantee a high rating. In fact, it is in the Worst 3 category. In contrast ingredients like Niacinamide or products that target dark spots and the skin's texture seem to be dominating the Top 3.

Now lets narrow this down to gain a deeper understanding of the metric's correlations. This will give us a deeper understanding of whether or not the high ratings are linked to customer engagement and the relationship of these metrics could be a strength or weakness in a company's product. 

```{r}
# Scatterplot: Rating vs Reviews
ggplot(combined, aes(x = reviews, y = rating)) +
  geom_point(alpha = 0.6, color = "blue") +
  geom_smooth(method = "lm", se = FALSE, color = "red") +
  labs(
    title = "Relationship Between Reviews and Rating",
    x = "Number of Reviews",
    y = "Rating"
  ) +
  theme_minimal()

# Scatterplot: Rating vs Loves Count
ggplot(combined, aes(x = loves_count, y = rating)) +
  geom_point(alpha = 0.6, color = "green") +
  geom_smooth(method = "lm", se = FALSE, color = "red") +
  labs(
    title = "Relationship Between Loves Count and Rating",
    x = "Loves Count",
    y = "Rating"
  ) +
  theme_minimal()

# Scatterplot: Loves Count vs Reviews
ggplot(combined, aes(x = reviews, y = loves_count)) +
  geom_point(alpha = 0.6, color = "purple") +
  geom_smooth(method = "lm", se = FALSE, color = "red") +
  labs(
    title = "Relationship Between Reviews and Loves Count",
    x = "Number of Reviews",
    y = "Loves Count"
  ) +
  theme_minimal()


```
The first graph, showing the relationship between reviews and ratings, also 
exhibits a weak positive correlation. Products with more reviews tend to have 
slightly higher ratings, but most ratings remain in the 4 to 5 range, regardless
of review count. Reviews highlight engagement and visibility, while ratings 
reflect customer satisfaction. The clustering of ratings indicates that while 
customer satisfaction is generally high, reviews are a more dynamic metric for 
measuring product visibility and interaction.

The second graph, focusing on the relationship between loves count and ratings, 
reveals a weak positive correlation. While products with higher loves tend to 
have slightly higher ratings, the relationship is not strong, and ratings 
cluster around 4 to 5 regardless of popularity. Loves count measures customer 
loyalty and emotional appeal, while ratings assess satisfaction. The lack of a 
strong correlation suggests that a product's popularity does not always 
translate directly into higher satisfaction. This emphasizes the importance of 
analyzing these metrics together to capture a product’s full performance.

The third graph, examining the relationship between reviews and loves count,
shows a strong positive correlation, indicating that products with higher 
reviews also tend to have higher loves. This suggests that customer interaction 
through reviews aligns closely with emotional resonance and loyalty, as measured
by loves count. Loves count reflects a product's desirability and brand appeal, 
while reviews showcase customer engagement and feedback. Products that perform 
well in both metrics demonstrate strong marketing and customer alignment.

Quantitatively, the correlation between reviews and loves is significantly 
stronger compared to the relationships between reviews and ratings or ratings 
and loves. This suggests that products with higher reviews are not only engaging
customers but are also more likely to be loved, making the combination of these 
metrics a holistic indicator of product performance.

```{r}
correlations <- combined %>%
  group_by(secondary_category) %>%
  summarize(
    corr_reviews_rating = cor(reviews, rating, use = "complete.obs"),  
    corr_rating_loves = cor(rating, loves_count, use = "complete.obs"),
    corr_reviews_loves = cor(reviews, loves_count, use = "complete.obs"), 
    .groups = "drop"
  )

correlations


```

This table reveals something interesting. The worst 3 products tend to have a higher average review and love count compared to the top 3 products which have fewer reviews and love counts. This suggests that high visibility and customer engagement (throguh reviews) dont always translate to positive sentiments or high ratings. instead these metrics could indicate polarizing products that attract significant attention but also receive miexed or negative feedback. However, the top 3 products,despite having fewer reviews and loves maintain a higher average rating which could mean they deliver consistent qualtiy and satisifcation to a samller but liekly more loyal audience. This observation could enforce the idea taht quantity of engagement throguh reviews or loves does not necessairly equate to a quality of sentiment through ratings. Understanding this dynamic could help a brand focus on improving their underperformign product with high visibility to turn into loyalty and satsifcation. But why might a low rated product have such low love counts you might ask? This could be due to either strong marketing, or a brand's reputation, a subset of users who strongly dislike and those who like, historical popularity that has declined, brand loyalty, or just simply an intent to purchase but never did.

```{r}
summary_metrics <- combined %>%
  group_by(category) %>%
  summarize(
    avg_rating = mean(rating, na.rm = TRUE),
    avg_reviews = mean(reviews, na.rm = TRUE),
    avg_loves = mean(loves_count, na.rm = TRUE),
    .groups = "drop"
  )

print(summary_metrics)

ggplot(summary_metrics, aes(x = category, y = avg_rating, fill = category)) +
  geom_bar(stat = "identity") +
  labs(
    title = "Average Rating Across Categories (Top 3 vs Worst 3)",
    x = "Category",
    y = "Average Rating"
  ) +
  theme_minimal() 

ggplot(summary_metrics, aes(x = category, y = avg_reviews, fill = category)) +
  geom_bar(stat = "identity") +
  labs(
    title = "Average Reviews Across Categories (Top 3 vs Worst 3)",
    x = "Category",
    y = "Average Reviews"
  ) +
  theme_minimal() 

ggplot(summary_metrics, aes(x = category, y = avg_loves, fill = category)) +
  geom_bar(stat = "identity") +
  labs(
    title = "Average Loves Across Categories (Top 3 vs Worst 3)",
    x = "Category",
    y = "Average Loves"
  ) +
  theme_minimal() 

```


```{r}
#use primary category now for graph simplification
combined_with_primary <- combined %>%
  left_join(product_info %>% select(product_id, primary_category), by = "product_id")

#remove "mini size" irrelevant and skews info
broad_category_summary <- combined_with_primary %>%
  filter(primary_category != "Mini Size") %>% 
  mutate(
    broad_category = ifelse(is.na(primary_category) | primary_category == "", "Other", primary_category)  
  ) %>%
  group_by(broad_category, category) %>%  
  summarize(
    avg_rating = mean(rating, na.rm = TRUE),   
    avg_reviews = mean(reviews, na.rm = TRUE), 
    avg_loves = mean(loves_count, na.rm = TRUE), 
    .groups = "drop"
  )

#avg rating
ggplot(broad_category_summary, aes(x = reorder(broad_category, avg_rating), y = avg_rating, fill = category)) +
  geom_bar(stat = "identity", position = "dodge") +
  coord_flip() +
  labs(
    title = "Average Ratings by Primary Categories (Top 3 vs Worst 3)",
    x = "Primary Category",
    y = "Average Rating",
    fill = "Category"
  ) +
  theme_minimal()

#avg review
ggplot(broad_category_summary, aes(x = reorder(broad_category, avg_reviews), y = avg_reviews, fill = category)) +
  geom_bar(stat = "identity", position = "dodge") +
  coord_flip() +
  labs(
    title = "Average Reviews by Primary Categories (Top 3 vs Worst 3)",
    x = "Primary Category",
    y = "Average Reviews",
    fill = "Category"
  ) +
  theme_minimal()

#avg love count
ggplot(broad_category_summary, aes(x = reorder(broad_category, avg_loves), y = avg_loves, fill = category)) +
  geom_bar(stat = "identity", position = "dodge") +
  coord_flip() +
  labs(
    title = "Average Loves Count by Primary Categories (Top 3 vs Worst 3)",
    x = "Primary Category",
    y = "Average Loves Count",
    fill = "Category"
  ) +
  theme_minimal()

```
To better understand the performance of products across primary categories, we 
analyzed the data based on average ratings, reviews, and loves counts for the 
top 3 and bottom 3 products in each category. Mini-sized products were excluded 
from this analysis to ensure the results focused on the full-sized offerings, as
mini sizes might skew the findings without offering unique insights. This 
breakdown allows for a clearer understanding of customer engagement, 
satisfaction, and loyalty within each category.

The results reveal several key trends. In skincare, the worst 3 products 
surprisingly received higher reviews and loves counts compared to the top 3 
products. This suggests that skincare products in the worst-performing group 
still garner significant customer interest but likely fail to meet expectations 
in terms of quality or effectiveness. This pattern could point to either high 
demand for these products, driven by strong marketing or desirable features, or 
simply a broader audience that ended up being dissatisfied with the results. 
This highlights the importance of investigating variability within this 
category, as high engagement does not necessarily translate into high 
satisfaction.

Conversely, Tools & Brushes emerge as a standout category, excelling across all 
metrics for the top 3 products. This indicates consistent quality and customer 
satisfaction, where tools and brushes meet or exceed expectations. The strong 
performance in reviews and loves counts reflects a loyal customer base and 
suggests that functionality and quality play critical roles in shaping consumer 
opinions in this category. Tools & Brushes demonstrate that delivering reliable,
high-quality products can lead to success across all performance metrics.

Other categories, such as Makeup, Hair, and Bath & Body, reveal more mixed 
patterns, where engagement levels and satisfaction are not always aligned. For 
example, some products in these categories have high reviews or loves counts but
still fall into the bottom 3 due to lower ratings. This suggests that polarizing
preferences or niche appeals might influence customer perceptions, with some 
consumers finding these products highly satisfactory while others are left 
disappointed. Fragrance, in particular, shows lower overall reviews and loves 
counts for both top and bottom performers, reflecting its position as a more 
specialized category that caters to a specific audience.

These findings emphasize the complex dynamics between engagement, loyalty, and 
satisfaction. While reviews and loves counts provide valuable insights into 
customer interest and emotional connection to products, they do not always 
correlate directly with ratings, which reflect overall satisfaction and quality.
Skincare, for instance, underscores the need for brands to explore why high 
customer interest does not always lead to high satisfaction, while the 
consistent performance of Tools & Brushes highlights the importance of quality 
in driving positive outcomes.

To gain a more nuanced understanding of these trends, it is essential to dive 
deeper into customer review sentiments. This would help uncover the specific 
reasons behind high engagement but low satisfaction in certain categories and 
identify potential areas for improvement. Additionally, leveraging the success 
of consistently high-performing categories like Tools & Brushes could provide 
valuable insights into how other categories might enhance their appeal and 
performance. Understanding these patterns will allow brands to refine their 
strategies, improve under performing products, and build on the strengths of 
their most successful offerings.

